### 2013-03-23 drafts/grammar/grammar.tex
Added section about Bayesian intepretation of E-C.

\usepackage{amsmath}
\usepackage{ amssymb }
\usepackage{graphicx}
\usepackage{syntax}
\usepackage{tikz}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}

@@

\begin{document}

In this document, I will describe a stochastic grammar over typed
combinatory logic expressions.

@@

>>>>

\usepackage{amsmath}
\usepackage{ amssymb }
\usepackage{graphicx}
\usepackage{syntax}
\usepackage{tikz}
\usepackage{tikz}
\usetikzlibrary{bayesnet}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}

@@

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\author{Eyal Dechter}
\begin{document}
\maketitle
\date

@@

\section{E-C as MAP inference in a probabilistic graphical model.}
The motivation of the E-C algorithm is that a learning agent should
learn a library of useful program fragments that represents knowledge
about the kind the programs that are likely to solve problems in a
given domain. One way to frame this is as a probabilistic generative
model of the tasks that an agent encounters
(Figure~\ref{fig:generativeModel}).

\begin{figure}
\begin{tikzpicture}[->, circle]

\node[latent] (theta) {$\theta$} ;
\node[latent, below=of theta] (G) {G} ;
\node[latent, below=of G]     (e) {$e_i$} ;
\node[obs, below=of e]     (t) {$t_i$} ;

\edge {theta} {G};
\edge {G} {e};
\edge {e} {t};

\plate {plate} {
  (e)(t)}{$i=1, \dots, N$};

\end{tikzpicture}
\caption{Generative model of program induction
  tasks. \label{fig:generativeModel}}
\end{figure}

Here each task $t_i$ is interpreted as a noisy version of the
underlying program $e_i$, and there is an associated likelihood
function $p_{t|e}(t|e) \propto f(t(e)) $ where $f$ is a monotonically
decreasing function. Given a task-specific loss function $t(\cdot)$, a
natural likelihood is $p_{t|e}(t|e) \propto \exp^{-t(e)}$. A set of
programs $\{e_i\}$ is drawn from a grammar $G$, which functions as the
subroutine library. Our prior expectations about $G$ are encoded in
$\theta$ via the distribution $p_{G|\theta}$.

The goal of the learner is the find the library $G$ and programs
$\{e_i\}$ that maximize $P(\{e_i\}, G | \{t_i\}, \theta)$. That is,

\begin{align}
G^*, \{e^*_i\} &= \argmax_{G, \{e_i\} }
   P(\{e_i\}, G | \{t_i\}, \theta)\\
&= \argmax_{G, \{e_i\} } P(\{t_i\} | \{e_i\})
  P(\{e_i\} | G) P(G | \theta)\\
&= \argmax_{G, \{e_i\} }
  \sum_i \left( 
  \log{p_{t|e}(t_i|e_i)}+ 
  \log{p_{e|G}(e_i|G)}
  \right ) +
  \log{p_{G|\theta}(G|\theta)}
\label{eq:objective}
\end{align}

The main difficulty in this optimization is that finding any candidate
program $e_i$ that provides a significant value for the first term is
difficult. If we new $G$, we could sample programs from it as a way to
guide searching over the $e_i$. This suggests the following iterative
heuristic search algorithm. At iteration $j$, let $G^{(j)}$ be the
current best value of $G$.

\begin{enumerate}
\item \textbf{Exploration step: } Let $F$ be a set of $T$ samples
  i.i.d from $p_{e|G}(\cdot | G^{(j)})$.
\item \textbf{Compression step: } 
$$G^{(j+1)}, \{e^{(j+1)}_i\} = 
\argmax_{G, \{e_i | e_i \in F \}}
\sum_i \left( 
\log{p_{t|e}(t_i|e_i)}+ 
\log{p_{e|G}(e_i|G)}
\right ) +
\log{p_{G|\theta}(G|\theta)}
$$
\end{enumerate}

%% We can break this joint maximization into an iteration of
%% coordinate-wise maximization steps. First we fix $G$ and maximize the
%% programs, then we fix the programs and maximize $G$, repeating until
%% we are no longer improving out objective. This gives us the following
%% iterative algorithm, roughly corresponding to the indicated steps of
%% the E-C algorithm:

%% \begin{enumerate}
%% \item \textbf{Compression step} $$G^{(j+1)} = 
%%   \argmax_{G} \sum_i \left( 
%%   \log{p_{e|G}(e^{(j)}_i|G)}
%%   \right ) +
%%   \log{p_{G|\theta}(G|\theta)}$$
%% \item \textbf{Exploration step} $$\{e^{(j+1)}_i\} = 
%%   \argmax_{\{e_i\}}
%%   \sum_i \left( 
%%   \log{p_{t|e}(t_i|e_i)}+ 
%%   \log{p_{e|G}(e_i|G^{(j)})}
%%   \right ) $$
%% \end{enumerate}
